---
title: "Exploratory Data Analysis"
format: html
editor: visual
---

```{r, echo=FALSE}
source("../.Rprofile")
here()
data <- read.csv(here('Output/finaldata.csv'))

head(data)
```

```{r, echo = FALSE}
tail(data)

```

```{r, echo=FALSE}
data %>%
  summary()

data <- data %>%
  mutate(conflict = as.character(conflict),
         drought = as.character(drought),
         earthquake = as.character(earthquake))
```

-   mat mor has a lot of NAs

```{r}
na_count <- colSums(is.na(data))

print(data.frame(Column = names(na_count), NA_Count = na_count))

```

-   mat mortality has large number of NAs (426)
-   all other mortality has 40 NAs

```{r}
data %>% count(ISO)
```

-   all countries have 20 entries

## Ploting distribution of Continuous Data

-   variables we are interested in are: all mortality variables
-   conflict variables

### plotting mat mortality

```{r}
 data %>%
  ggplot(aes(x = Year, y = mat.mor, group = ISO))+
  geom_line(color = '#eb9ad3', alpha = 7/10)+
  scale_y_log10()+
  ggtitle('maternal mortality')
```

-   hard to see any distinct patterns

```{r}
 data %>%
  ggplot(aes(x = Year, y = infant.mor, group = ISO))+
  geom_line(color = '#89CFF0', alpha = 7/10)+
  scale_y_log10()+
  ggtitle('infant mortality')
```

-   looks like overall it's going down

```{r}
 data %>%
  ggplot(aes(x = Year, y = neo.mor, group = ISO))+
  geom_line(color = '#cbbdf2')+
  scale_y_log10()+
  ggtitle('neonatal mortality')
```

-   looks to also be decreasing slightly

```{r}
 data %>%
  ggplot(aes(x = Year, y = under5.mor, group = ISO))+
  geom_line(color = '#a1efcf')+
  scale_y_log10()+
  ggtitle('Under 5 mortality')
```

-   looks to be decreasing over years

### How many countries have conflict?

```{r}
data %>%
  ggplot(aes(conflict))+
  geom_bar()+
  ggtitle('conflict')
```

## Which Countries have conflict? 
```{r}
data %>%
  group_by(country_name, conflict) %>%
  summarize(sumcon = sum(conflict == 1)) %>%
  dplyr::filter(conflict == 1) %>%
  arrange(sumcon) %>%
  print(data)

```
- most countries only had a few years where conflict occured 

### group mortality by conflict 
```{r}
data %>%
  ggplot(aes(x = conflict, y = mat.mor, fill = conflict))+
  geom_violin()+
  ggtitle('maternal mortality grouped by country conflict')
```
- countries with no conflict have lower rates of maternal mortality 

```{r}
data %>%
  ggplot(aes(x = conflict, y = infant.mor, fill = conflict))+
  geom_violin()+
  ggtitle('infant mortality grouped by country conflict')
```

```{r}
data %>%
  ggplot(aes(x = conflict, y = neo.mor, fill = conflict))+
  geom_violin()+
  ggtitle('neonatal mortality grouped by country conflict')
```
```{r}
data %>%
  ggplot(aes(x = conflict, y = under5.mor, fill = conflict))+
  geom_violin()+
  ggtitle('infant mortality grouped by country conflict')
```





